Continuous-Time Markov Chain with Alarms analysis and comparaison of tools \thanks{Supported by organization x.}

### 2018-07-12 Report/report/paper.tex
change in the paper for my institution

\subsection{Time distribution}
The model we use are probabilistic continuous time model, which means that the time passed in a state will depend on a time distribution. A time distribution can be depicted by its cumulative distribution function (CDF) $F : \mathbb{R} \rightarrow[0,1]$, $F(t)$ is the probability that the event will happend before the time $t$. The exponential distribution of rate $\lambda$ has a CDF $F(t) = 1 - \exp(-\lambda t)$, it is important to keep in mind that a rate refers to the exponential distribution. The Dirac distribution of timeout $\tau$ has a CDF $F(t) = \chi_{[\tau,+\infty[}(t) = 
\left\{
	\begin{array}{l}
		0 \text{ if } t < \tau\\
		1\text{ else}
	\end{array}
\right.$. In this paper, time distribution which are not exponential will be called non-Markovian distribution.
\subsection{CTMC}
A continuous-time Markov chain (CTMC) is a triple $C = (S,Q,s_{in})$, where 
\begin{itemize}
	\item[$\bullet$] S is a finite set of states
	\item[$\bullet$] $Q : S \times S \rightarrow \mathbb{R}_{\geq 0}$ is a matrix of rates such that $\sum_{s' \in S} Q(s,s')  > 0$ for each $s \in S$
	\item[$\bullet$] $s_{in} \in S$ is an initial state
\end{itemize} 
We define the exit rate of state $s$ as $\lambda_s = \sum_{s' \in S} Q(s,s')$

@@

To obtain analysis on ACTMC, we use PH fitting to create CTMC from ACTMC, and analyse those CTMC Then we deduce results on the ACTMC.  The reason is that Prism or Storm can use CTMC and not ACTMC. Figure \ref{fig:simple_actmc} and figure \ref{fig:simple_ctmc} are a toy example of the PH fitting of a simple ACTMC $\mathcal{A} = (\{A,B\},
\begin{bmatrix}
	1       & 3 \\
	2       & 1 
\end{bmatrix}
,A,\{A\},
\begin{bmatrix}
1       & 0
\end{bmatrix}
,\{d\})$ with constant parameter $k$. The created CTMC is $\mathcal{C} = (\{A,B,1,2,...,k-1,B\},
\begin{bmatrix}
0       & \lambda_{1,k} 	& 0		& 0		& ...		& 0\\
0       & 0 	& \lambda_{2,k}		& 0		& ...		& 0\\
0       & 0 	& 0		& \lambda_{3,k}		& ...		& 0\\
0       & 0 	& 0		& 0		& ...		& 0\\
...       & ... 	& ...		& ...		& ...		& \lambda_{k,k}\\
0       & 0 	& 0		& 0		& 0		& 0
\end{bmatrix}
,A)$

@@

		\node[state] (s0)                    {$A$};
		\node[state]         (sn) [right of=s0]       {$B$};
		\path[->, dashed]
		(s0) edge 	node {$d$} (sn);

		\end{tikzpicture}
		\caption{Model of a simple ACTMC $\mathcal{A}$, $d$ is a non-Markovian distribution.}
		\label{fig:simple_actmc}
	\end{figure}

@@

		\path[->]  
		(s0) edge	node {$\lambda_{k,1}$} (s1);
		\path[->]

		[->]
		(s1) edge	node {$\lambda_{k,2}$} (s2);
		\path[->]

		(s2) edge   node {$\lambda_{k,3}$} (s4);
		\path[->]

		(s4) edge   node {$\lambda_{k,k-1}$} (sw);
		\path[->]

		(sw) edge   node {$\lambda_{k,k}$} (sn);
		\path[->];
		\end{tikzpicture}
		\caption{Model of a CTMC obtain by PH fitting of the ACTMC figure \ref{fig:simple_actmc}.}
		\label{fig:simple_ctmc}
	\end{figure}

>>>>

\subsection{Time distribution}
The model we use are probabilistic continuous time model, which means that the time passed in a state will depend on a time distribution. A time distribution can be depicted by its cumulative distribution function (CDF) $F : \mathbb{R} \rightarrow[0,1]$, $F(t)$ is the probability that the event will happend before the time $t$. The exponential distribution of rate $\lambda$ has a CDF $F(t) = 1 - \exp(-\lambda t)$, it is important to keep in mind that a rate refers to the exponential distribution. The Dirac distribution of timeout $\tau$ has a CDF $F(t) = \chi_{[\tau,+\infty[}(t) = 
\left\{
	\begin{array}{l}
		0 \text{ if } t < \tau\\
		1\text{ else}
	\end{array}
\right.$. In this paper, time distribution which are not exponential will be called non-Markovian distribution.
\subsection{CTMC}
A continuous-time Markov chain (CTMC) is a triple $C = (S,Q,s_{in})$, where 
\begin{itemize}
	\item[$\bullet$] S is a finite set of states
	\item[$\bullet$] $Q : S \times S \rightarrow \mathbb{R}_{\geq 0}$ is a matrix of rates such that $\sum_{s' \in S} Q(s,s')  > 0$ for each $s \in S$
	\item[$\bullet$] $s_{in} \in S$ is an initial state
\end{itemize} 
$Q(s,s')$ is the rate of the transition from $s$ to $s'$.
We define the exit rate of state $s$ as $\lambda_s = \sum_{s' \in S} Q(s,s')$

@@

To obtain analysis on ACTMC, we use PH fitting to create CTMC from ACTMC, and analyse those CTMC Then we deduce results on the ACTMC.  The reason is that Prism or Storm can use CTMC and not ACTMC. Figure \ref{fig:simple_actmc} and figure \ref{fig:simple_ctmc} are a toy example of the PH fitting of a simple ACTMC $\mathcal{A} = (\{A,B\},
\begin{bmatrix}
	0.5       & 0 \\
	2       & 1 
\end{bmatrix}
,A,\{A\},
\begin{bmatrix}
1       & 0
\end{bmatrix}
,\{d\})$ with constant parameter $k$. The created CTMC is $\mathcal{C} = (\{A,B,1,2,...,k-1,B\},
\begin{bmatrix}
0       & \lambda_{1,k} 	& 0		& 0		& ...		& 0\\
0       & 0 	& \lambda_{2,k}		& 0		& ...		& 0\\
0       & 0 	& 0		& \lambda_{3,k}		& ...		& 0\\
0       & 0 	& 0		& 0		& ...		& 0\\
...       & ... 	& ...		& ...		& ...		& \lambda_{k,k}\\
0       & 0 	& 0		& 0		& 0		& 0
\end{bmatrix}
,A)$

@@

		\node[state] (s0)                    {$A$};
		\node[state]         (sn) [right of=s0]       {$B$};
		\path[->, dashed]
		(s0) edge 	node {$d$} (sn);
		\path[->]
		(s0) edge [loop left]	node {0.5} (s0)

		(sn) edge [bend left]	node {2} (s0)
		edge [loop right]	node {1} (sn);			
		\end{tikzpicture}
		\caption{Model of a simple ACTMC $\mathcal{A}$, $d$ is a non-Markovian distribution.}
		\label{fig:simple_actmc}
	\end{figure}

@@

		\path[->]  
		(s0) edge	node {$\lambda_{k,1}$} (s1)
			edge [loop above]	node {0.5} (s0);
		\path[->]

		[->]
		(s1) edge	node {$\lambda_{k,2}$} (s2)
			edge [loop above]	node {0.5} (s1);
		\path[->]

		(s2) edge   node {$\lambda_{k,3}$} (s4)
			edge [loop above]	node {0.5} (s2);
		\path[->]

		(s4) edge   node {$\lambda_{k,k-1}$} (sw)
			edge [loop above]	node {0.5} (s4);
		\path[->]

		(sw) edge   node {$\lambda_{k,k}$} (sn)
			edge [loop above]	node {0.5} (sw);
		\path[->]

		(sn) edge [bend left]	node {2} (s0)
		edge [loop right]	node {1} (sn);
		\end{tikzpicture}
		\caption{Model of a CTMC obtain by PH fitting of the ACTMC figure \ref{fig:simple_actmc}.}
		\label{fig:simple_ctmc}
	\end{figure}
